---
title: "Palmer penguins"
author: "David"
date: "26/11/2021"
output: html_document
---

```{r libraries setup, echo=FALSE, message=FALSE, warning=FALSE}
if (!require('ggplot2')) install.packages('ggplot2'); library('ggplot2')
if (!require('palmerpenguins')) install.packages('palmerpenguins'); library('palmerpenguins')
if (!require('knitr')) install.packages('knitr'); library('knitr')
if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
if (!require('tidyr')) install.packages('tidyr'); library('tidyr')
```

## Introduction:

In this notebook are reports and observations about penguins from Palmer archipelago.

In detail I will report:

* The species that inhabit Palmer archipelago, organized by each Island
* The relation between the length of flippers and body mass of the different penguins species
* The different sizes of flipper for each specie and the most common flipper size per specie
* The difference in weight for each specie from female to male and the average body mass for each

In the meantime if you would prefer you can find more information about Palmer penguins by clicking [here](https://en.wikipedia.org/wiki/Palmer_Station) .

## Let's start 

### The species that inhabit Palmer archipelago, organized by each Island

```{r species per island table, echo=FALSE, message=FALSE, warning=FALSE}
inhabited_islands <- penguins %>% group_by(species) %>% summarize(unique(island))
inhabited_islands_df <- data.frame(inhabited_islands)
colnames(inhabited_islands_df) <- c("Species", "Island of Residence")
kable(inhabited_islands_df)
```

### The relation between the length of flippers and body mass of the different penguins species

```{r bill length vs body mass, echo=FALSE, message=FALSE, warning=FALSE}
penguins %>% drop_na() %>% ggplot() + 
  geom_jitter(mapping=aes(y=body_mass_g, x=flipper_length_mm, color=species)) + 
  xlab("Length of flippers") + ylab("Body mass") + labs(color="Species") + 
  annotate(geom = "curve", x = 215, y = 3500, xend = 222, yend = 4500, 
    curvature = .3, arrow = arrow(length = unit(2, "mm"))) + 
  annotate(geom= "text", x=215, y=3250, label="The biggest specie")
```

### The different sizes of flipper for each specie and the most common flipper size per specie

```{r sizes of flippers, echo=FALSE, message=FALSE, warning=FALSE}
   messages_df <- data.frame(label = c("190 mm", "196 mm", "217 mm"),
                            species = c("Adelie", "Chinstrap", "Gentoo"),
                            xcord = c(210, 215, 197),
                            ycord= c(12.5, 12.5, 12.5)
  )
  flippers_mean <- penguins %>% drop_na() %>% group_by(species) %>% summarize(mean_of_flippers = mean(flipper_length_mm))
  
  penguins %>% drop_na() %>% ggplot() + geom_bar(mapping=aes(x=flipper_length_mm, fill=species)) + facet_wrap(~species) + 
geom_vline(data=flippers_mean, mapping=aes(xintercept=mean_of_flippers), color = "black", linetype="dashed") +
labs(y="Number of Penguins", x="Length of the flippers (mm)", fill="Species") + 
geom_text(data= messages_df, mapping=aes(x = xcord, y = ycord, label = label, fontface="italic"), color="black")
```

###  The difference in weight for each specie from female to male and the average body mass for each

```{R female to male weight by specie, echo=FALSE, message=FALSE, warning=FALSE}
penguins_mean_weight_df <- data.frame(species = c("Adelie", "Adelie", "Chinstrap", "Chinstrap", "Gentoo", "Gentoo"),
                                      sex = c("female", "male", "female", "male", "female", "male"),
                                      mean_weight_label = c("3369 g", "4043 g", "3527 g", "3939 g", "4680 g", "5485 g"),
                                      xcord = c(4000, 4500, 4000, 4500, 5000, 5000),
                                      ycord = c(5, 5, 5, 5, 5, 5),
                                      adjustment= c(0, 0, 0, 0, 0, 1 )
                                     )

penguins_mean_weight_line <- penguins %>% drop_na() %>% group_by(species, sex) %>% summarize(mean_weight = mean(body_mass_g))

penguins %>%  drop_na() %>% ggplot() + geom_bar(mapping=aes(x=body_mass_g, fill=species)) + 
  facet_wrap(vars(species, sex)) + labs(y="Number of penguins", x="Body mass (g)", fill="Species") + 
  geom_vline(data=penguins_mean_weight_line, mapping=aes(xintercept=mean_weight), color="black", linetype="dashed") + geom_text(data=penguins_mean_weight_df, mapping=aes(x = xcord, y= ycord, label = mean_weight_label, hjust = adjustment), color="black", fontface="italic")
```